A Formal Model of Checked C

### 2022-11-28 sandp23/background.tex
Minor

\myparagraph{Checked Pointer Types}
\checkedc{} introduces three varieties of \emph{checked pointer}:
\begin{itemize}
\item \code{ptr<}$T$\code{>} types a pointer that is either null or
  points to a single object of type $T$.
\item \code{array_ptr<}$T$\code{>} types a pointer that is either null
  or points to an array of $T$ objects. The array width is defined
  by a \emph{bounds} expression, discussed below.
\item \code{nt_array_ptr<}$T$\code{>} is like
  \code{array_ptr<}$T$\code{>} except that the bounds expression
  defines the \emph{minimum} array width---additional objects may
  be available past the upper bound, up to a null terminator.
\end{itemize}
A bounds expression used with the latter two pointer types has three
forms:
\begin{itemize}
\item \code{count(}$e$\code{)} where $e$ defines the array's
  length. Thus, if pointer $p$ has bounds \code{count(n)} then the
  accessible memory is in the range $[p,p+$\code{n}$]$. Bounds
  expression $e$ must be side-effect free and may only refer to
  variables whose addresses are not taken, or adjacent \code{struct}
  fields.
\item \code{byte_count(}$e$\code{)} is like \code{count}, but
  expresses arithmetic using bytes, no objects; i.e.,
  \code{count(}$e$\code{)} used for \code{array_ptr<}$T$\code{>} is
  equivalent to \code{byte_count(}$e\times\texttt{sizeof}(T)$\code{)}
\item \code{bounds(}$e_l$,$e_h$\code{)} where $e_l$ and $e_h$ are
  pointers that bound the accessible region $[e_l,e_h)$ (the
  expressions are similarly restricted). Bounds
  \code{count(}$e$\code{)} is shorthand for
  \code{bounds(}$p, p + e$\code{)}. This most general form of bounds
  expression is useful for supporting pointer arithmetic.
\end{itemize}
  Dropping the bounds expression on an \code{nt_array_ptr} is equivalent
  to the bounds being \code{count(0)}.

>>>>

\noindent
\myparagraph{Checked Pointer Types}
\checkedc{} introduces three varieties of \emph{checked pointer}:
\begin{itemize}
\item \code{ptr<}$T$\code{>} ($\ptr$) types a pointer that is either null or
  points to a single object of type $T$.
\item \code{array_ptr<}$T$\code{>} ($\arr$) types a pointer that is either null
  or points to an array of $T$ objects. The array width is defined
  by a \emph{bounds} expression, discussed below.
\item \code{nt_array_ptr<}$T$\code{>} ($\ntarr$) is like
  \code{array_ptr<}$T$\code{>} except that the bounds expression
  defines the \emph{minimum} array width---additional objects may
  be available past the upper bound, up to a null terminator.
\end{itemize}
Both $\arr$ and $\ntarr$ pointers have an associated bounds which defines the
range of memory referenced by the pointer.
The three different ways to declare bounds and the corresponding memory range is:
\begin{footnotesize}
\begin{tabular}{ll}
\arrT{|$T$|} \inlinecode{p: count(|$n$|)}
  &
$[\inlinecode{p}, \inlinecode{p}+\inlinecode{sizeof}(T) \times n) $ \\
%$[\inlinecode{p}, )$ \\
\arrT{|$T$|} \inlinecode{p: byte_count(|$b$|)}

  &
    $[\code{p}, \code{p}+b)$ \\

\arrT{|$T$|} \inlinecode{p: bounds(|$x, y$|)}

  &
    $[x, y)    $\\  
\end{tabular}
\end{footnotesize}
The bounds can be declared for $\ntarr$ as well, but the memory range can extend further to the right,
until a~\code{NULL} terminator is reached (\ie the~\code{NULL} is not within the bounds).

\subsection{Ensuring Memory Safety}

%A bounds expression used with the latter two pointer types has three
%forms:
%\begin{itemize}
%\item \code{count(}$e$\code{)} where $e$ defines the array's
%  length. Thus, if pointer $p$ has bounds \code{count(n)} then the
%  accessible memory is in the range $[p,p+$\code{n}$]$. Bounds
%  expression $e$ must be side-effect free and may only refer to
%  variables whose addresses are not taken, or adjacent \code{struct}
%  fields.
%\item \code{byte_count(}$e$\code{)} is like \code{count}, but
%  expresses arithmetic using bytes, no objects; i.e.,
%  \code{count(}$e$\code{)} used for \code{array_ptr<}$T$\code{>} is
%  equivalent to \code{byte_count(}$e\times\texttt{sizeof}(T)$\code{)}
%\item \code{bounds(}$e_l$,$e_h$\code{)} where $e_l$ and $e_h$ are
%  pointers that bound the accessible region $[e_l,e_h)$ (the
%  expressions are similarly restricted). Bounds
%  \code{count(}$e$\code{)} is shorthand for
%  \code{bounds(}$p, p + e$\code{)}. This most general form of bounds
%  expression is useful for supporting pointer arithmetic.
%\end{itemize}
%
%\begin{footnotesize}
%\begin{tabular}{ll}
%\arrT{|$T$|} \inlinecode{p: count(|$n$|)}
%  &
%$[\inlinecode{p}, \inlinecode{p}+\inlinecode{sizeof}(T) \times n) $ \\
%%$[\inlinecode{p}, )$ \\
%\arrT{|$T$|} \inlinecode{p: byte_count(|$b$|)}
%
%  &
%    $[\code{p}, \code{p}+b)$ \\
%
%
%\arrT{|$T$|} \inlinecode{p: bounds(|$x, y$|)}
%
%  &
%    $[x, y)    $\\  
%\end{tabular}
%\end{footnotesize}
%
%  Dropping the bounds expression on an \code{nt_array_ptr} is equivalent
%  to the bounds being \code{count(0)}.
